---
title: "1. Preprocessing S2 images with level-2A"
author: "Jean-Baptiste Féret"
date: "`r Sys.Date()`"
output:
  html_vignette:
    number_sections: true
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 2
vignette: >
  \usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{1. Preprocessing S2 images with level-2A}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---
  
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```

This tutorial aims at illustrating how to access and preprocess Sentinel-2 images acquired at Level-2A with the R package `preprocS2`. 


# Identifying and downloading Sentinel-2 images

In the present situation, we are interested in downloading a unique Sentinel-2 image with no cloud cover for an illustration of the package. 
We selected the Barrax area in Spain, as it is a well-known study area for application using remote sensing for crop monitoring. 

First, we used [Sentinel-hub](https://apps.sentinel-hub.com/eo-browser/?zoom=11&lat=39.09836&lng=-2.01462&themeId=DEFAULT-THEME&visualizationUrl=https%3A%2F%2Fservices.sentinel-hub.com%2Fogc%2Fwms%2Fbd86bcc0-f318-402b-a145-015f85b9427e&datasetId=S2L2A&fromTime=2021-05-13T00%3A00%3A00.000Z&toTime=2021-05-13T23%3A59%3A59.999Z&layerId=1_TRUE_COLOR) in order to identify an acquisition corresponding to our requirement. 

Once the S2 tile and the date of acquisition were identified, we could proceed to the image download. 
We first defined a polygon corresponding to our study area, named 'Study_area_Barrax.shp' and saved on our computer. 

We then used the R package [`sen2r`](https://sen2r.ranghetti.info/) in order to identify S2 product available for a specific date of acquisition, and download the corresponding Sentinel-2 SAFE archive. Here, level-2A images (atmospherically corrected with Sen2Cor) can be downloaded directly from [SciHub](https://scihub.copernicus.eu) with sen2r.


```{r download_S2_image}
library(sen2r)
# define date of S2 acquisition
dateAcq <- '2021-05-13'
time_window <- as.Date(c(dateAcq, dateAcq))
# define path for study area
path_vector <- '../DATA/Study_area_Barrax.shp'
# get product name corresponding to Barrax 
list_safe <- s2_list(spatial_extent = sf::st_read(dsn = path_vector), time_interval = time_window)
# define output directory where SAFE zipfile is stored
DirWrite <- '../DATA/S2_Images'
s2_download(list_safe, outdir=DirWrite)

```

# Preprocess S2 image

Once the SAFE S2 image is downloaded, the R package [`preprocS2`](https://jbferet.gitlab.io/preprocS2/index.html) is used to stack individual bands, crop to required extent defined by the vector file, produce a binary cloud mask and save in the desired raster format.

The function `extract_from_S2_L2A` performs cropping and resampling to 10m if needed, and provides a stars object as output, along with location for metadata and image full name.

```{r preprocess_S2}
library(preprocS2)
##____________________________________________________________________##
##        Define where data is stored and where to write results      ##
##--------------------------------------------------------------------##
# define raster path
Path_S2 <- file.path(DirWrite,list.files(DirWrite,pattern = '.SAFE'))
# Result directory
result_path <- '../RESULTS'
dir.create(path = result_path,showWarnings = FALSE,recursive = TRUE)

##____________________________________________________________________##
##                  Extract, resample & stack data                    ##
##--------------------------------------------------------------------##
# define resolution
resolution <- 10
# define source of data
S2source <- 'SAFE'
S2obj <- preprocS2::extract_from_S2_L2A(Path_dir_S2 = Path_S2,
                                        path_vector = path_vector,
                                        S2source = S2source,
                                        resolution = resolution)

# update shapefile if needed (reprojection)
path_vector <- S2obj$path_vector

```

# Write reflectance, cloud mask and metadata

Once the image is read, it can be written as stacked raster file, and stored with the binary cloud mask and the metadata file. 

```{r write_raster}
# create specific result directory corresponding to granule name
results_site_path <- file.path(result_path,basename(S2obj$S2_Bands$GRANULE))
dir.create(path = results_site_path,showWarnings = FALSE,recursive = TRUE)
##____________________________________________________________________##
##                        Write CLOUD MASK                            ##
##--------------------------------------------------------------------##
# directory for cloud mask
Cloud_path <- file.path(results_site_path,'CloudMask')
dir.create(path = Cloud_path,showWarnings = FALSE,recursive = TRUE)
# Filename for cloud mask
cloudmasks <- preprocS2::save_cloud_s2(S2_stars = S2obj$S2_Stack,
                                       Cloud_path = Cloud_path,
                                       S2source = S2source, SaveRaw = T)
##____________________________________________________________________##
##                        Write REFLECTANCE                           ##
##--------------------------------------------------------------------##
# directory for Reflectance
Refl_dir <- file.path(results_site_path,'Reflectance')
dir.create(path = Refl_dir,showWarnings = FALSE,recursive = TRUE)
# filename for Reflectance
Refl_path <- file.path(Refl_dir,paste(basename(S2obj$S2_Bands$GRANULE),'_Refl',sep = ''))

# Save Reflectance file as ENVI image with BIL interleaves
tile_S2 <- substring(strsplit(basename(S2obj$S2_Bands$GRANULE),'_')[[1]][2],2)
dateAcq_S2 <- as.Date(substring(strsplit(basename(S2obj$S2_Bands$GRANULE),'_')[[1]][4],1,8),format="%Y%m%d")
preprocS2::save_reflectance_s2(S2_stars = S2obj$S2_Stack, Refl_path = Refl_path,
                               S2Sat = NULL, tile_S2 = tile_S2, dateAcq_S2 = dateAcq_S2,
                               Format = 'ENVI_BIL', datatype = 'Int16', MTD = S2obj$S2_Bands$metadata)
```


# Computation of spectral indices and biophysical variables

Once reflectance data is written as a raster file, the R package `prosail` can be used for the computation of various spectral indices and biophysical variables direcrtlty from a raster stack corresponding to Level-2A reflectance. 


