---
title: "1. Get Sentinel-2 imagery from Planetary STAC catalog"
author: "Jean-Baptiste Féret"
date: "`r Sys.Date()`"
output:
  html_vignette:
    number_sections: true
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 2
vignette: >
  \usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{1. Get Sentinel-2 imagery from Planetary STAC catalog}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```

This tutorial illustrates how to download Sentinel-2 data for a given bounding box with the R package `preprocS2`.

`preprocS2` is a basic wrapper of the [`rstac`](https://brazil-data-cube.github.io/rstac/) package. 
It aims at downloading and preparing Sentinel-2 images for further processings using R packages such as [`biodivMapR`](https://jbferet.github.io/biodivMapR/index.html) for spectral diversity mapping, [`prosail`](https://jbferet.gitlab.io/prosail/index.html) for the computation of vegetation biophysical variables, or [`spinr`](https://gitlab.com/jbferet/spinr) for the computation of vegetation biophysical variables or spectral indices.

# How to use preprocS2

`preprocS2` will download S2 data from a STAC catalog based on spatial and temporal information.
The default STAC collection is `sentinel-2-l2a` from the Planetary computer catalog.

Optionally, `preprocS2` can download the geometry of acquisition corresponding to the S2 acquisition.
Access to geometry of acquisition is provided by the package [`CDSE`](https://zivankaraman.github.io/CDSE/).

Please provide authentication for CDSE if you want to get geometry of acquisition by activating OAuth clients following [`this link`](https://shapps.dataspace.copernicus.eu/dashboard/#/account/settings).

## Download image corresponding to a bounding box

User can provide a bounding box as input to define the region of interest.
User is also requested to provide a valid date of acquisition The Sentinel-2 tile corresponding to the region of interest is identified based on the S2 tiling grid system.
User can either download the corresponding kml [`here`](https://sentiwiki.copernicus.eu/web/s2-products), or let `preprocS2` download it.
Access online is requested to download the tiling grid system.

```{r get s2 from bbox}
# library
library(preprocS2)
library(sf)

# 1- define input & output directories
output_dir <- './RESULTS_bbox'
dir.create(path = output_dir, showWarnings = F, recursive = T)
datetime <- as.Date('2024-10-05')

# 2- define area of interest
bbox <- sf::st_bbox(obj = c('xmin' = -54.1, 'ymin' = 5.0, 
                            'xmax' = -54.0, 'ymax' = 5.1))

# get tiling grid kml from https://sentiwiki.copernicus.eu/web/s2-products
path_S2tilinggrid <- 'Sentinel-2_tiling_grid.kml'

# 3- get S2 acquisition and geometry of acquisition for study area
# https://shapps.dataspace.copernicus.eu/dashboard/#/account/settings
id <- "sh-XXXXXXXX-XXXX-XXXX-XXXX-XXXXXXXXXXXX"
secret <- "XXXXXXXXXXXXXXXXXXXXXXXXXXXX"
authentication <- list('id' = id, 'pwd' = secret)

get_s2_raster(bbox = bbox, datetime = datetime, output_dir = output_dir, 
              path_S2tilinggrid = path_S2tilinggrid, siteName = 'FrenchGuiana',
              overwrite = F, geomAcq = T, authentication = authentication)
```

The output files are stored in the directory `output_dir`.
Different subdirectories are created.

-   `collections` includes the item collection corresponding to the scene of interest
-   `raster_samples` includes rasters corresponding to BOA reflectance, scene classification (SCL) and a binary mask identifying vegetated pixels after radiometric filters excluding shaded pixels, cloudy pixels and non-vegetated pixels
-   `geomAcq_S2` includes rasters corresponding to the geometry of acquisition (SAA, SZA, VAA, VZA)

## Download image corresponding to a vector file

Alternatively, user can provide the path for a vector file defining a region of interest.

```{r get s2 from vector}
# library
library(preprocS2)
library(sf)

# 1- define input & output directories
output_dir <- './RESULTS_vector'
dir.create(path = output_dir, showWarnings = F, recursive = T)
datetime <- as.Date('2024-10-05')

# 2- define area of interest
bbox <- sf::st_bbox(obj = c('xmin' = -54.1, 'ymin' = 5.0, 
                            'xmax' = -54.0, 'ymax' = 5.1))
aoi_path <- file.path(output_dir, 'S2_FrenchGuiana_aoi.GPKG')
sf::st_write(obj = bbox_to_poly(x = bbox), dsn = aoi_path, driver = 'GPKG', overwrite = T)
# get tiling grid kml from https://sentiwiki.copernicus.eu/web/s2-products
path_S2tilinggrid <- 'Sentinel-2_tiling_grid.kml'

# 3- get S2 acquisition and geometry of acquisition for study area
# https://shapps.dataspace.copernicus.eu/dashboard/#/account/settings
id <- "sh-XXXXXXXX-XXXX-XXXX-XXXX-XXXXXXXXXXXX"
secret <- "XXXXXXXXXXXXXXXXXXXXXXXXXXXX"
authentication <- list('id' = id, 'pwd' = secret)

get_s2_raster(aoi_path = aoi_path, datetime = datetime, output_dir = output_dir, 
              path_S2tilinggrid = path_S2tilinggrid, siteName = 'FrenchGuiana',
              overwrite = F, geomAcq = T, authentication = authentication)
```

## Additional functionalities

S2 images corresponding to large extents can be downloaded using aoi tiling and parallel download/process to speed it up.
This will be documented soon.

