---
title: "1. Get Sentinel-2 imagery from Planetary STAC catalog"
author: "Jean-Baptiste Féret"
date: "`r Sys.Date()`"
output:
  html_vignette:
    number_sections: true
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 2
vignette: >
  \usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{1. Get Sentinel-2 imagery from Planetary STAC catalog}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```

This tutorial illustrates how to download Sentinel-2 data for a given bounding 
box with the R package `preprocS2`.

`preprocS2` is a basic wrapper of the 
[`rstac`](https://brazil-data-cube.github.io/rstac/) package.
It aims at downloading and preparing Sentinel-2 images for further processings 
using R packages such as 
[`biodivMapR`](https://jbferet.github.io/biodivMapR/index.html) for spectral 
diversity mapping, [`prosail`](https://jbferet.gitlab.io/prosail/index.html) for 
the computation of vegetation biophysical variables, or 
[`spinr`](https://gitlab.com/jbferet/spinr) for the computation of spectral 
indices.

# How to use preprocS2

`preprocS2` will download S2 data from a STAC catalog based on spatial and 
temporal information.
The default STAC collection is `sentinel-2-l2a` from the 
[`Planetary computer`](https://planetarycomputer.microsoft.com/dataset/sentinel-2-l2a) catalog.

Optionally, `preprocS2` can download the geometry of acquisition corresponding 
to the S2 acquisition.
Access to geometry of acquisition is provided by the package 
[`CDSE`](https://zivankaraman.github.io/CDSE/).

Please provide authentication for CDSE if you want to get geometry of 
acquisition by activating OAuth clients following 
[`this link`](https://shapps.dataspace.copernicus.eu/dashboard/#/account/settings).

## Download S2 subset from a STAC catalog based on a bounding box

User can provide a bounding box as input to define the region of interest.
User is also requested to provide a valid date of acquisition.
The Sentinel-2 tile corresponding to the region of interest is identified based 
on the S2 tiling grid system.
User can either download the corresponding kml 
[`here`](https://sentiwiki.copernicus.eu/web/s2-products), or let `preprocS2` 
download it.
Access online is requested to download the tiling grid system.

```{r get s2 from bbox}
# library
library(preprocS2)
library(sf)

# 1- define input & output directories
output_dir <- './RESULTS_bbox'
dir.create(path = output_dir, showWarnings = F, recursive = T)
datetime <- as.Date('2024-10-05')

# 2- define area of interest
bbox <- sf::st_bbox(obj = c('xmin' = -54.1, 'ymin' = 5.0, 
                            'xmax' = -54.0, 'ymax' = 5.1))
sf::st_crs(x = bbox) <- 4326

# a CRS is required for the bbox. 
# downloaded Sentinel-2 image will be reprojected using this CRS
# make sure you define a CRS corresponding to S2 product if you do not want reprojection and asociated resampling

# get tiling grid kml from https://sentiwiki.copernicus.eu/web/s2-products
path_S2_tiling_grid <- 'Sentinel-2_tiling_grid.kml'

# 3- get S2 acquisition 
list_files <- get_s2_raster(bbox = bbox, datetime = datetime, 
                            output_dir = output_dir, 
                            path_S2_tiling_grid = path_S2_tiling_grid, 
                            site_name = 'FrenchGuiana',
                            overwrite = F, geom_acq = T)

# Sentinel-2 L2A reflectance
Refl_L2A <- list_files$Refl_L2A
# Sentinel-2 binary mask identifying vegetation, discarding clouds & shadows
vegetation_mask <- list_files$vegetation_mask
# Sentinel-2 mask from provider ( = SCL from ESA products)
SCL <- list_files$provider_mask
# Sentinel-2 geometry of acquisition (if requested from CDSE)
geometryAcquisition <- list_files$geometryAcquisition
```

The output files are stored in the directory `output_dir`.
Different subdirectories are created.

-   `collections` includes the item collection corresponding to the scene of interest
-   `raster_samples` includes rasters corresponding to BOA reflectance, scene classification (SCL) and a binary mask identifying vegetated pixels after radiometric filters excluding shaded pixels, cloudy pixels and non-vegetated pixels
-   `geom_acq_S2` includes rasters corresponding to the geometry of acquisition (SAA, SZA, VAA, VZA)

## Download S2 subset from a STAC catalog based on a vector footprint

Alternatively, user can provide the path for a vector file defining a region of interest.

```{r get s2 from vector}
# library
library(preprocS2)
library(sf)

# 1- define input & output directories
output_dir <- './RESULTS_vector'
dir.create(path = output_dir, showWarnings = F, recursive = T)
datetime <- as.Date('2024-10-05')

# 2- define area of interest
bbox <- sf::st_bbox(obj = c('xmin' = -54.1, 'ymin' = 5.0, 
                            'xmax' = -54.0, 'ymax' = 5.1))
sf::st_crs(x = bbox) <- 4326

# a CRS is required for the bbox. 
# downloaded Sentinel-2 image will be reprojected using this CRS
# make sure you define a CRS corresponding to S2 product if you do not want reprojection and asociated resampling

aoi_path <- file.path(output_dir, 'S2_FrenchGuiana_aoi.GPKG')
sf::st_write(obj = bbox_to_poly(x = bbox), dsn = aoi_path, 
             driver = 'GPKG', delete_layer = TRUE)
# get tiling grid kml from https://sentiwiki.copernicus.eu/web/s2-products
path_S2_tiling_grid <- 'Sentinel-2_tiling_grid.kml'

# 3- get S2 acquisition 
list_files <- get_s2_raster(aoi_path = aoi_path, datetime = datetime, 
                            output_dir = output_dir, 
                            path_S2_tiling_grid = path_S2_tiling_grid, 
                            site_name = 'FrenchGuiana',
                            overwrite = F, geom_acq = T)
# Sentinel-2 L2A reflectance
Refl_L2A <- list_files$Refl_L2A
# Sentinel-2 binary mask identifying vegetation, discarding clouds & shadows
vegetation_mask <- list_files$vegetation_mask
# Sentinel-2 mask from provider ( = SCL from ESA products)
SCL <- list_files$provider_mask
# Sentinel-2 geometry of acquisition (if requested from CDSE)
geometryAcquisition <- list_files$geometryAcquisition
```

## Extract S2 subset from a vector footprint and a SAFE S2 product

`preprocS2` can be used to crop and stack reflectance data from a __L2A SAFE__ archive provided by ESA.
First, download a SAFE file corresponding to your area of interest and unzipped it. 
The function `extract_from_S2_L2A` will then provide all information required to crop and stack the reflectance data. 
The L2A reflectance corresponding to the vector footprint is saved with `save_reflectance_s2`
The corresponding cloud mask is saved with `save_cloud_s2`. 
The code below illustrates the full process. 

```{r extract from SAFE}
library(preprocS2)
safe_path <- 'path/to/SAFE/S2A_MSIL2A_XXXXXXXXX.SAFE'
path_vector <- 'path/to/VectorFootprint'

output_dir <- '../S2_crop'
# define resolution
resolution <- 10
# define source of data
S2source <- 'SAFE'
S2obj <- extract_from_S2_L2A(Path_dir_S2 = safe_path,
                             path_vector = path_vector,
                             S2source = S2source,
                             resolution = resolution)


# create specific result directory corresponding to granule name
results_site_path <- file.path(output_dir,basename(S2obj$S2_Bands$GRANULE))
dir.create(path = results_site_path,showWarnings = FALSE,recursive = TRUE)
##____________________________________________________________________##
##                        Write CLOUD MASK                            ##
##--------------------------------------------------------------------##
# directory for cloud mask
Cloud_path <- file.path(results_site_path,'CloudMask')
dir.create(path = Cloud_path,showWarnings = FALSE,recursive = TRUE)
# Filename for cloud mask
cloudmasks <- save_cloud_s2(S2_stars = S2obj$S2_Stack,
                            Cloud_path = Cloud_path,
                            S2source = S2source, SaveRaw = T)
##____________________________________________________________________##
##                        Write REFLECTANCE                           ##
##--------------------------------------------------------------------##
# directory for Reflectance
Refl_dir <- file.path(results_site_path,'Reflectance')
dir.create(path = Refl_dir,showWarnings = FALSE,recursive = TRUE)
# filename for Reflectance
Refl_path <- file.path(Refl_dir,paste(basename(S2obj$S2_Bands$GRANULE),'_Refl',sep = ''))

# Save Reflectance file as ENVI image with BIL interleaves
# metadata files are important to account for offset applied on S2 L2A products 
tile_S2 <- preprocS2::get_tile(S2obj$S2_Bands$GRANULE)
dateAcq_S2 <- preprocS2::get_dateAcq(S2product = dirname(dirname(S2obj$S2_Bands$GRANULE)))
save_reflectance_s2(S2_stars = S2obj$S2_Stack, 
                    Refl_path = Refl_path,
                    S2Sat = NULL, 
                    tile_S2 = tile_S2, 
                    dateAcq_S2 = dateAcq_S2,
                    Format = 'ENVI', 
                    datatype = 'Int16', 
                    MTD = S2obj$S2_Bands$metadata, 
                    MTD_MSI = S2obj$S2_Bands$metadata_MSI)
```

# Go further and download larger regions

S2 images corresponding to large extents can be downloaded using aoi tiling and 
parallel download / process to speed it up.

Tutorial vignettes describing these additional functionalities are available 
[here](https://jbferet.github.io/biodivMapR/articles/biodivMapR_01.html).
