---
title: "1. Preprocessing S2 images with level-2A"
author: "Jean-Baptiste Féret"
date: "`r Sys.Date()`"
output:
  html_vignette:
    number_sections: true
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 2
vignette: >
  \usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{1. Preprocessing S2 images with level-2A}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---
  
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```

This tutorial aims at illustrating how to access and preprocess Sentinel-2 images acquired at Level-2A with the R package `preprocS2`. 


# Identifying and downloading Sentinel-2 images

This tutorial aims at illustrating how to use preprocS2 to download and prepare Sentinel-2 images for further processings using R packages such as [`biodivMapR`](https://jbferet.github.io/biodivMapR/index.html) for spectral diversity mapping, or [`prosail`](https://jbferet.gitlab.io/prosail/index.html) for the computation of vegetation biophysical variables or spectral indices

We are interested in downloading a unique Sentinel-2 image with no cloud cover for an illustration of the package. 
We selected the Barrax area in Spain, as it is a well-known study area for application using remote sensing for crop monitoring. 

### Download vector files used to identify study area

We first define a data directory (modify following your personal taste) and download a shapefile corresponding to our study area.
The following code allows downloading the zipfile containing the vector data, and unzipping it in the same directory as the one including the raster data.

```{r download vector data}
# library
library(zip)
# define data directory
Path_Data <- '../../01_DATA'
dir.create(Path_Data,showWarnings = F, recursive = T)
# name zip file including plots located on the tile
destzip <- file.path(Path_Data,'Study_area_Barrax.zip',fsep = '\\')
# url for the zip file
url <- 'https://gitlab.com/jbferet/myshareddata/-/raw/master/preprocS2_Example/Study_area_Barrax.zip'
download.file(url = url, destfile = destzip)
destunz <- file.path(Path_Data,'Study_area_Barrax',fsep = '\\')
unzip(zipfile = destzip,exdir = destunz)
```



Then, we use [Sentinel-hub](https://apps.sentinel-hub.com/eo-browser/?zoom=11&lat=39.09836&lng=-2.01462&themeId=DEFAULT-THEME&visualizationUrl=https%3A%2F%2Fservices.sentinel-hub.com%2Fogc%2Fwms%2Fbd86bcc0-f318-402b-a145-015f85b9427e&datasetId=S2L2A&fromTime=2021-05-13T00%3A00%3A00.000Z&toTime=2021-05-13T23%3A59%3A59.999Z&layerId=1_TRUE_COLOR) in order to identify an acquisition corresponding to our requirement. 

Once the S2 tile and the date of acquisition are identified, we can proceed to the image download. 

We then used the R package [`sen2r`](https://sen2r.ranghetti.info/) in order to identify S2 product available for a specific date of acquisition, and download the corresponding Sentinel-2 SAFE archive. Here, level-2A images (atmospherically corrected with Sen2Cor) can be downloaded directly from [SciHub](https://scihub.copernicus.eu) with sen2r.

```{r download_S2_image_sen2r}
library(sen2r)
# define date of S2 acquisition
dateAcq <- '2022-05-28'
time_window <- as.Date(c(dateAcq, dateAcq))
# define path for study area
path_vector <- '../../01_DATA/Study_area_Barrax/Study_area_Barrax.shp'
# get product name corresponding to Barrax 
list_safe <- s2_list(spatial_extent = sf::st_read(dsn = path_vector), time_interval = time_window)
# define output directory where SAFE zipfile is stored
DirWrite <- '../../01_DATA/S2_Images'
dir.create(DirWrite,showWarnings = F, recursive = T)
s2_download(list_safe, outdir=DirWrite)

# define raster path
Path_S2 <- file.path(DirWrite,list.files(DirWrite,pattern = '.SAFE'))

```


You can also use the function `get_S2_L2A_Image` from `preprocS2` in order to download this image. This function uses the `s2_download` from `sen2r`. If a google cloud accounbt is parameterized in `sen2r`, you can alternatively download S2 images from it by setting `GoogleCloud = TRUE`.

```{r download_S2_image_preprocS2}
library(preprocS2)
# define date of S2 acquisition
dateAcq <- '2022-05-28'
# define path for study area
path_vector <- '../../01_DATA/Study_area_Barrax/Study_area_Barrax.shp'
# define output directory where SAFE zipfile is stored
DirWrite <- '../../01_DATA/S2_Images'
dir.create(DirWrite,showWarnings = F, recursive = T)

Path_S2 <- get_S2_L2A_Image(l2a_path = DirWrite, 
                            spatial_extent = path_vector, 
                            dateAcq = dateAcq,
                            DeleteL1C = TRUE, 
                            Sen2Cor = TRUE,
                            GoogleCloud = TRUE)


```


# Preprocess S2 image

Once the SAFE S2 image is downloaded, the R package [`preprocS2`](https://jbferet.gitlab.io/preprocS2/index.html) is used to stack individual bands, crop to required extent defined by the vector file, produce a binary cloud mask and save in the desired raster format.

The function `extract_from_S2_L2A` performs cropping and resampling to 10m if needed, and provides a stars object as output, along with file path for metadata and image full name.

```{r preprocess_S2}
library(preprocS2)
##____________________________________________________________________##
##        Define where data is stored and where to write results      ##
##--------------------------------------------------------------------##
# Result directory
result_path <- '../../03_RESULTS'
dir.create(path = result_path,showWarnings = FALSE,recursive = TRUE)

##____________________________________________________________________##
##                  Extract, resample & stack data                    ##
##--------------------------------------------------------------------##
# define resolution
resolution <- 10
# define source of data
S2source <- 'SAFE'
S2obj <- preprocS2::extract_from_S2_L2A(Path_dir_S2 = Path_S2,
                                        path_vector = path_vector,
                                        S2source = S2source,
                                        resolution = resolution)

# update shapefile if needed (reprojection)
path_vector <- S2obj$path_vector

```

# Write reflectance, cloud mask and metadata

Once the image is read, it can be written as stacked raster file, and stored with the binary cloud mask and the metadata file.
Since 2021, S2 images are provided with an [offset](https://sentinels.copernicus.eu/web/sentinel/-/copernicus-sentinel-2-major-products-upgrade-upcoming) in order to prevent from negative radiometric values. 
In order to account for this offset when writing reflectance file (and get actual reflectance values with a multiplying factor of 10 000), the S2 metadata files (MTD and MTD_MSI) are required. The path for these files is provised as an output of `extract_from_S2_L2A`.

```{r write_raster}
# create specific result directory corresponding to granule name
results_site_path <- file.path(result_path,basename(S2obj$S2_Bands$GRANULE))
dir.create(path = results_site_path,showWarnings = FALSE,recursive = TRUE)
##____________________________________________________________________##
##                        Write CLOUD MASK                            ##
##--------------------------------------------------------------------##
# directory for cloud mask
Cloud_path <- file.path(results_site_path,'CloudMask')
dir.create(path = Cloud_path,showWarnings = FALSE,recursive = TRUE)
# Filename for cloud mask
cloudmasks <- preprocS2::save_cloud_s2(S2_stars = S2obj$S2_Stack,
                                       Cloud_path = Cloud_path,
                                       S2source = S2source, SaveRaw = T)
##____________________________________________________________________##
##                        Write REFLECTANCE                           ##
##--------------------------------------------------------------------##
# directory for Reflectance
Refl_dir <- file.path(results_site_path,'Reflectance')
dir.create(path = Refl_dir,showWarnings = FALSE,recursive = TRUE)
# filename for Reflectance
Refl_path <- file.path(Refl_dir,paste(basename(S2obj$S2_Bands$GRANULE),'_Refl',sep = ''))

# Save Reflectance file as ENVI image with BIL interleaves
# metadata files are important to account for offset applied on S2 L2A products 
tile_S2 <- get_tile(S2obj$S2_Bands$GRANULE)
dateAcq_S2 <- get_date(S2obj$S2_Bands$GRANULE)
preprocS2::save_reflectance_s2(S2_stars = S2obj$S2_Stack, 
                               Refl_path = Refl_path,
                               S2Sat = NULL, 
                               tile_S2 = tile_S2, 
                               dateAcq_S2 = dateAcq_S2,
                               Format = 'ENVI', 
                               datatype = 'Int16', 
                               MTD = S2obj$S2_Bands$metadata, 
                               MTD_MSI = S2obj$S2_Bands$metadata_MSI)
```


# Computation of spectral indices and biophysical variables

Once reflectance data is written as a raster file, the R package `prosail` can be used for the computation of various spectral indices and biophysical variables direcrtlty from a raster stack corresponding to Level-2A reflectance. 


